---
title: "What makes region attractive for worker?"
author: "Yuga Hikida"
date: Sys.Date()
fontsize: 12pt
output: 
  beamer_presentation:
  theme: "Boadilla"
  innertheme: "circles"
header-includes:
  - \definecolor{UBCblue}{rgb}{0.04706, 0.13725, 0.26667} % UBC Blue (primary)
  - \usecolortheme[named=UBCblue]{structure}
  - \setbeamercolor{frametitle}{fg=white,bg=UBCblue}
  - \setbeamercolor{titlelike}{parent=structure,fg=UBCblue}
  - \setbeamertemplate {footline}{\quad\hfill\insertframenumber\strut\quad}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include = FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(bayesplot)
library(brms)
library(patchwork)
library(latex2exp)
```


```{r include = FALSE}
theme_set(theme_minimal() +
theme(axis.text = element_text(size=15),
      axis.title = element_text(size = 20),
      legend.title = element_text(size=20),
      legend.text = element_text(size = 20),
      strip.text = element_text(size=20),
      plot.title = element_text(size = 20)
      ))
```


## Background

* Determinant of migration
 + Labour market condition
   - (net) wage
   - Unemployment rate
 + Amenities 
   - Public goods
   - facilities

* Emprically analyse what makes region attractive for worker.
 + Focus on changes in emplopyees in each region over time.
 + $y_t := 100 \times \{ \log (\text{employed}_t)  - \log (\text{employed}_{t-1}) \}$
 + Period: 2016 - 2021
 + All data obtained from <https://www.inkar.de/>
 
 
## Literature review


 
## Explanatory variables



## Basic panel data model

For region $i=1,...,N$, time $t=1,..,T$:

$$
y_{it} = \alpha + \beta^T X_{it} + u_{it} \\
u_{it} = \mu_i + \nu_{it} \\
\mu_i \overset{\mathrm{iid}}{\sim} (0, \sigma_\mu) \\
\nu_{it} \overset{\mathrm{iid}}{\sim} (0, \sigma_\nu)
$$

where $X_{it} = [x_{it1}, ..., x_{itk}]^T$ and $\beta = [\beta_1,...,\beta_k]^T$.

$\mu_i$ is a group specific effect ("random effect") differs across region 
which is modelled independently constant over time.


## Spatially correlated group specific effect

$$
y_{it} = \alpha + \beta^T X_{it} + \epsilon_{it} \\
\epsilon_{it} = \rho \sum_{j=1, j \neq i}^{N} w_{ij} \epsilon_{ij} + \nu_{it}\\
\epsilon_{it} \overset{\mathrm{iid}}{\sim} (0, \sigma_\epsilon) \\
\nu_{it} \overset{\mathrm{iid}}{\sim} (0, \sigma_\nu)
$$




