---
title: "What makes region attractive for worker?"
author: "Yuga Hikida"
institute: "TU Dortmund University"
fontsize: 11pt
format: 
  beamer:
    incremental: false
    theme: "metropolis"
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r include = FALSE}
library(tidyr)
library(dplyr)
library(mice)
library(ggplot2)
library(plm)
library(splm)
library(sf)
library(spdep)
library(sf)
library(viridis)
library(jtools)
library(patchwork)
```


```{r include = FALSE}
theme_set(theme_minimal() +
theme(axis.text = element_text(size=15),
      axis.title = element_text(size = 20),
      legend.title = element_text(size=20),
      legend.text = element_text(size = 20),
      strip.text = element_text(size=20),
      plot.title = element_text(size = 20)
      ))
```

```{r include = FALSE}
set.seed(1)
d_wide <- read.csv("inkar/data.csv")

# Data pre-processing
col_eng <- c("code_num", "spatial_unit", "year", "employed", "unemployment_rate",
             "land_price", "employees_academic_degree", "employees_no_qualification",
             "hospital_bed", "prop_foreigners", "waste_amount", "gdp")

d <- d_wide %>% 
  slice(-1) %>%
  select(- Aggregat) %>%
  pivot_longer(
    cols = - c(Kennziffer, Raumeinheit),
    names_to = c(".value", "year"),
    names_pattern = "(.+?)\\.?(\\d*)$") %>%
  replace(.=="", "0") %>%
  setNames(col_eng) %>%
  mutate(year = recode(year,
                       "0" = "2015",
                       "1" = "2016",
                       "2" = "2017",
                       "3" = "2018",
                       "4" = "2019",
                       "5" = "2020",
                       "6" = "2021"),
         d_employed = 100 * (log(employed) - log(dplyr::lag(employed))))%>%
  filter(year != "2015")
  

# Impute data
imputed <- mice(d, method="cart")
d <- complete(imputed)
anyNA(d)

# transform some variables
d <- d %>% mutate(ln_gdp = log(gdp),
                  sc_land_price = land_price / 1000,
                  sc_waste_amount = waste_amount / 10000,
                  year = as.factor(year),
                  spatial_unit = as.factor(spatial_unit)) %>% as_tibble()

## with shape data
shapefile <- st_read("vg2500_krs/vg2500_krs.shp")
shapefile$ARS <- as.numeric(shapefile$ARS)

d_geo <- merge(shapefile, d, by.x = "ARS", by.y = "code_num", all.y = TRUE)
```


## Background

* Determinant of migration of worker [@moretti2011local]
  + Labour market condition
    - net wage, unemployment rate...
  + Amenities 
    - public goods, facilities...

* **Empirically analyse what makes region attractive for worker.**

* Previous studies: migration decision from various factors
from both of the channels [@buch2014makes; @hakim2022amenities].
  + No single strong determinants of migration.
 
## Data
 * Focus on changes in emplopyees in each region over time.
 * $y_t := 100 \times \{ \log (\text{employed}_t)  - \log (\text{employed}_{t-1}) \}$
 * Period: 2016 - 2021
 * All data is obtained from <https://www.inkar.de/>
 
```{r}
#| label: map-employed
#| fig-cap: "Changes in #employed (%)"
#| fig-subcap:
#|   - "2016"
#|   - "2021"
#| layout-ncol: 2

d_geo %>%
  filter(year == "2016") %>%
  mutate(d_employed_q = cut(d_employed, breaks = quantile(d_employed, probs = c(0, 0.20, 0.5, 0.8, 1)), 
                            include.lowest = TRUE)) %>%
  ggplot() +
  geom_sf(aes(fill = d_employed_q)) +
  scale_fill_viridis_d(option = "magma", direction = -1)

d_geo %>%
  filter(year == "2021") %>%
  mutate(d_employed_q = cut(d_employed, breaks = quantile(d_employed, probs = c(0, 0.20, 0.5, 0.8, 1)), 
                            include.lowest = TRUE)) %>%
  ggplot() +
  geom_sf(aes(fill = d_employed_q)) +
  scale_fill_viridis_d(option = "magma", direction = -1)
```


## Modalities

- `splm` package [@millo2009splm] is used for all the estimation.
- Maximum likelihood method is used.
- Statistical testing: 5% significance level.
 
## Explanatory variables

\scriptsize
\begin{table}[]
\begin{tabular}{@{}ll@{}}
\toprule
Variable name & description \\ \midrule
GDP per employed & GDP in €1,000 per employed person\\
Unemployement rate & Share of unemployed people in \% \\
Land price &  Average purchase values for building land in € per m2 \\
Proportion of foreigners  & Proportion of foreigners to residents in \% \\
Amount of waste & Amount of waste disposed of or treated per inhabitant in kg \\
Hospital beds & Hospital beds per 1,000 inhabitants \\
\toprule
\end{tabular}
\end{table}
\normalsize

- Some variables are transformed such that scale of all the variables are 
roughly same.
- Denoted as $X$ in the following slides.

## Basic panel data model

For region $i=1,...,N$, time $t=1,..,T$:

\begin{align*}
  y_{it} &= \alpha + \beta^T X_{it} + u_{it} \\
  u_{it} &= \mu_i + \nu_{it} \\
  \mu_i &\overset{\mathrm{iid}}{\sim} (0, \sigma_\mu) \\
  \nu_{it} &\overset{\mathrm{iid}}{\sim} (0, \sigma_\nu)
\end{align*}

where $X_{it} = [x_{it1}, ..., x_{itK}]^T$ and $\beta = [\beta_1,...,\beta_K]^T$.

$\mu_i$ is a group specific effect ("random effect") differs across region 
which is modelled independently constant over time.

```{r include = FALSE}
# delete row in shapefile which is not recorded in main data
setdiff(unique(shapefile$ARS), unique(d$code_num))
shapefile <- shapefile %>% filter(ARS != 16056)
  
# panel data with group specific effect
d <- d %>% select(- code_num)

f <- d_employed ~ ln_gdp + unemployment_rate + sc_land_price + prop_foreigners +
  sc_waste_amount + hospital_bed
  
fit1 <- plm(f, data = d, model = "random")
```

## Result

```{r fig.width=5, fig.height=3}
plot_summs(fit1, robost  = TRUE)
```

- Counter-intuitive result from economic theory
- Potential model misspecification?


## Adding autoregressive error term

\begin{align*}
  y_{it} &= \alpha + \beta^T X_{it} + u_{it} \\
  u_{it} &= \mu_i + \lambda \sum_{j=1, j \neq i}^{N} w_{ij} u_{jt} + \nu_{it}\\
  \mu_i &\overset{\mathrm{iid}}{\sim} (0, \sigma_\mu) \\
  \nu_{it} &\overset{\mathrm{iid}}{\sim} (0, \sigma_\nu)
\end{align*}

where $|\lambda| < 1$ is spatial autoregressive parameters and $w_{ij}$ is the 
$i,j$ element of $N \times N$ spatial weight matrix.

Now error term is decomposed into three parts: group specific part, 
autoregressive part and idiosyncratic part.

## Result: Autoregressive error model
- LM test under the null hypothesis $\lambda = 0$ suggests existence of spatial autocorrelation $\rightarrow$ estimate model with autoregressive error.

```{r include = FALSE}
# panel data with spacially correlated group specific effect

## creating weight matrix
krs_shp_mtrx <- sf::st_read("vg2500_krs") %>% filter(ARS != 16056)
krs_mtrx <- poly2nb(krs_shp_mtrx)
W <- nb2listw(krs_mtrx, style="W")

test1 <- bsktest(x = f, data = d, listw = W, test = "CLMlambda")
## adding special error term
fit2 <- spml(f, data = d, listw = W, model = "random", lag = FALSE)

fit2_vis <- fit1
fit2_vis$coefficients <- fit2$coefficients; fit2_vis$vcov <- fit2$vcov
```

```{r fig.width=5, fig.height=2.7}
plot_summs(fit2_vis, robost  = TRUE)
```

## Spatial lag of the dependent variable

\begin{align*}
  y_{it} &= \alpha + \rho \sum_{j=1, j \neq i}^{N} w_{ij} y_{jt} + \beta^T X_{it} + u_{it} \\
  u_{it} &= \mu_i + \lambda \sum_{j=1, j \neq i}^{N} w_{ij} u_{jt} + \nu_{it}\\
  \mu_i &\overset{\mathrm{iid}}{\sim} (0, \sigma_\mu) \\
  \nu_{it} &\overset{\mathrm{iid}}{\sim} (0, \sigma_\nu)
\end{align*}

- $|\rho| < 1$ is the effect of spatially lagged dependent variable.
- Consider so called direct impact and indirect effect (explained later).

## Result: Spatial lag model + spatially correlated error

```{r include = FALSE}
fit3 <- spml(f, data = d, listw = W, model = "random", lag = TRUE)
fit3_vis <- fit1
fit3_vis$coefficients <- fit3$coefficients; fit3_vis$vcov <- fit3$vcov

plot_lambda_conf_int <- function(lambda, vcov_lambda, crit){
  df <- tibble(var_name = "lagged_y",
                   Estimate = lambda,
                   lower = lambda - crit * sqrt(vcov_lambda),
                   upper = lambda + crit * sqrt(vcov_lambda))
  ggplot(df, aes(x = Estimate, y = var_name)) +
    geom_point(size = 3, shape = 1, color = "4AB7FC") +
    geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0, colour= "#4AB7FC", size = 1) +
    geom_vline(aes(xintercept = 0), colour = "black", linetype="dotted") +
    theme_minimal() +
    theme(axis.title.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x=element_blank())
}
```

```{r fig.width=5, fig.height=3}
plot_lambda_conf_int(fit3$arcoef, fit3$vcov.arcoef, 1.96) / plot_summs(fit3_vis, robost  = TRUE) + plot_layout(height  = c(1, 13))
```

- Difficult to interpret directly (next slide) 
- Statistical significance of lagged term implies potential model 
misspecification of other two models.

## Direct and indirect effect
Focus on single time. Matrix form of the model is given by
\begin{align*}
y &= \alpha + \rho W y + X \beta + u\\
y &= (I -  \rho W)^{-1} X \beta + (I -  \rho W)^{-1} (\alpha + u)\\
\tilde{y} :&= (I -  \rho W)^{-1} X \beta \\
&= \sum_{k=1}^{K} (I -  \rho W)^{-1} I \beta_k X_k
:= \sum_{k=1}^{K} S_k(W) X_k
\end{align*}
where $X_k = [x_{1},...,x_{Nk}]^T$.

- Direct effect = impact of $x_{ki}$ on $y_{i}$
  + Diagonal element of $S_k(W)$.
- Indirect effect = impact of $x_{kj}, j \neq i$ on $y_{i}$
  + Off-diagonal element of $S_k(W)$. "Spatial spillover".
- Average over region / time to obtain summary measures [@ruttenauer2024spatial].

## Result: Direct and indirect effect

```{r include = FALSE}
impac <- summary(impacts(fit3, listw = W, time = length(unique(d$year))))
```


```{r fig.width=7, fig.height=4}
#| label: two-effects
#| fig-cap: "Direct and indirect effect"

direct <- impac$direct_sum$quantiles %>%
  as.data.frame() %>%
  tibble() %>%
  setNames(c("lower", "nouse_1", "mean", "nouse_2", "upper")) %>%
  mutate(var_names = rownames(impac$direct_sum$quantiles)) %>%
  mutate(effect = "direct") %>%
  select(var_names, effect, lower, mean, upper) %>%
  pivot_longer(cols = "var_names")

indirect <- impac$indirect_sum$quantiles %>%
  as.data.frame() %>%
  tibble() %>%
  setNames(c("lower", "nouse_1", "mean", "nouse_2", "upper")) %>%
  mutate(var_names = rownames(impac$direct_sum$quantiles)) %>%
  mutate(effect = "indirect") %>%
  select(var_names, effect, lower, mean, upper) %>%
  pivot_longer(cols = "var_names")

combind <- bind_rows(direct, indirect)

ggplot(combind, aes(factor(value, levels = rev(attr(terms(f), "term.labels"))),
                    y = mean, ymin = lower, ymax = upper, color=effect, shape = effect)) + 
         geom_pointrange(position = position_dodge(width = 0.25), size=0.9) +
  geom_hline(yintercept = 0, colour = "black", linetype="dotted") + 
  ylab(NULL) + 
  xlab(NULL) + 
  coord_flip()
```


## Economic interpretation
- Positive direct effect / negative spillover in GDP per capita.
  + Worker move to region with high GDP.
- Negative direct effect / positive spillover in unemployment rate.
  + Worker move out of region with high employment rate.
- Positive direct effect / negative spillover in land price.
  + People move to region with high rent? 
  + Could be due to amenities associated with high rent.
  + Urban area, more amusement facilities etc.


## References
::: {#refs}
:::

## Appendix: LM test
\scriptsize
```{r}
test1 <- bsktest(x = f, data = d, listw = W, test = "CLMlambda")
test1
```


